---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Analysis of Personal Movement Using Activity Monitoring Devices - Week 2 Assignment 

## 1.Code for reading in the dataset and/or processing the data

```{r}
data<-read.csv("activity.csv")
head(data)
```

## 2.Histogram of the total number of steps taken each day


```{r, fig.width=4, fig.height=4,echo=TRUE}
dailysteps<-aggregate(steps ~ date,data, sum,na.rm=TRUE)
hist(dailysteps$steps,col="red", main="Histogram of Daily Steps")
```


## 3.Mean and median number of steps taken each day

```{r, echo=TRUE}
mean(dailysteps$steps,na.rm=TRUE)
median(dailysteps$steps,na.rm=TRUE)
```


## 4. Time series plot of the average number of steps taken
```{r, fig.width=8, fig.height=4, echo=TRUE}
timesteps<-aggregate(steps ~ interval,data, mean,na.rm=TRUE)
plot(steps ~ interval,timesteps, type="l" )
```


## 5. The 5-minute interval that, on average, contains the maximum number of steps
```{r, echo=TRUE}
timesteps$interval[timesteps$steps == max(timesteps$steps)]
```


## 6.Code to describe and show a strategy for imputing missing data
Missings value will be replaced by interval by the average of the equivalent available interval
```{r, echo=TRUE}
sum(is.na(data$steps))
```

## 7.create a new dataset that is equal to the original but with the missing data filled in. 
```{r,  fig.width=8, fig.height=4,echo=TRUE}
data2<-data
data2<-merge(data2,timesteps,by="interval")
data2$steps<-data2$steps.x
data2$steps[is.na(data2$steps.x)] <- data2$steps.y[is.na(data2$steps.x)]

dailysteps2<-aggregate(steps ~ date,data2, sum)
hist(dailysteps2$steps,col="yellow", main="Histogram of Daily Steps with NA missings value replaced")
```

## 8.Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r, fig.width=8, fig.height=4, echo=TRUE}
data2$date <- as.Date(data2$date, format = "%Y-%m-%d")
data2$day<-as.factor(weekdays(data2$date))

data2$weekday <- ifelse((data2$day %in% c("Saturday","Sunday")), "Weekend", "Weekdays") 

weekdayssteps<- aggregate(steps ~ weekday + interval,data2, mean)

library(lattice)
xyplot(steps ~  interval | weekday, data = weekdayssteps, layout = c(1,2), type ="l", ylab="Number of Steps - Weekdays vs weekend")
```